Growth Dynamics of the Independent Gametophytes of *Pleurosoriopsis makinoi*

### 2019-06-20 ms.Rmd
Get rmd to render to pdf
- Use knitr::include_graphics for external images.
- Use specific commit of gt to avoid latex bugs
- Use xelatex engine
- Add page breaks

---
title: "Growth Dynamics of the Independent Gametophytes of *Pleurosoriopsis makinoi* (Polypodiaceae)"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

@@

@@

```{r fig-1, fig.width=7, fig.height=8, fig.cap = "Fig. 1. The quadrats for measurement of gametophyte cover of *Pleurosoriopsis makinoi* in Unazawa Valley, Okutama-machi, Tokyo, Japan. a. A full view of the site with four quadrat positions (q1, q2, q3 and q4). b. A 10 cm × 10 cm quadrat frame."}

magick::image_read_pdf(here("ms/Fig1_quadrats.pdf"))

@@

```{r fig-2, fig.width=8, fig.height=7, fig.cap = "Fig. 2. Gametophytes of *Pleurosoriopsis makinoi* collected in Unazawa Valley, Okutama-machi, Tokyo, Japan. a. A typical ribbon-like gametophyte with gemmae on the apical part (arrowhead), collected in August. b. Gemmae (arrowheads) on a gametophyte collected in November. c. A gametophyte with archegonia (arrowheads), collected in October. d. Archegonia. e. A gametophyte with antheridia (arrowhead) collected in December. f. An antheridium. Scale bars: 2 mm for a, 200 μm for b and d, 1 mm for c and e, 100 μm for f."}

magick::image_read_pdf(here("ms/Fig2_gemma_sexualorgans.pdf"))

```{r fig-3, fig.width=8, fig.height=7, fig.cap="Fig. 3. Change in growth of *Pleurosoriopsis makinoi* gametophytes observed from May 2009 to May 2014. a. Cover (cm^2^) of gametophytes in four 10 × 10 cm quadrats. b. Mean number of gemmae per individual. c. Mean gemma length (μm). For b and c, *n* = 10 individuals per observation; grey bars show SD. Shaded bars show alternating years."}
#' Set x-limits manually since the automatically set limits will vary
#' when we subset the data.
start_date <- combined_morph %>%
  pull(date) %>%
  min

@@

```{r fig-4, fig.width=8, fig.height=7, fig.cap = "Fig. 4. Change in microclimate at site with *Pleurosoriopsis makinoi* independent gametophytes (Okutama) and sporophytes (Uratakao) from 2013 to 2014. a--b. Daily light integral (DLI) (mol m^-2 day^-1). c--d. Daily minimum relative humidity (%). e--f. Daily mean temperature (°C). Shaded bars show alternating years. Data collected at Okutama from September 22, 2012 to July 22, 2014 and at Uratakao from August 20, 2013 to November 22, 2014. Some data missing due to equipment malfunction (see Results)."}

@@

```{r fig-5, evaluate = FALSE, fig.width=7, fig.height=8, fig.cap="Fig. 5. Density plots showing difference between microclimate at *Pleurosoriopsis makinoi* independent gametophyte site (Okutama) and sporophyte site (Uratakao). Significant differences (paired t-test) shown by asterisks: \\*\\*\\*, *p* < 0.001; \\*\\*, *p* < 0.01; \\*, *p* < 0.05.  Y-axis is relative probability density; curves scaled so that total probability of each scales to 1. Vertical lines shown medians."}
#' Extract p-values for plotting, and put into data frame
#' with xvals specifying season and yvals specifying height
#' to plot asterisks.
p_vals <-
  t_test_results %>%
  mutate(pval = case_when(
    pval < 0.0001 ~ "***",
    pval < 0.001 ~ "**",
    pval < 0.05 ~ "*",
    TRUE ~ ""
  )) %>%
  select(season, var, pval) %>%
  spread(var, pval) %>%
  rename(
    par_total_pval = par_total,
    rh_min_pval = rh_min,
    temp_mean_pval = temp_mean
  )

@@

ggsave(glue::glue("fig-5_{Sys.Date()}.pdf"), width = 7, height = 8, units = "in")
```

```{r fig-6, fig.width=7, fig.height=8, fig.cap="Fig. 6. Relationships between growth of *Pleurosoriopsis makinoi* gametophytes (total cover, mean gemma number, and mean gemma length) and microclimate at the Okutama site. Microclimatic variables are monthly means of daily light integral (DLI) (mol m^-2^ day^-1^), daily minimum relative humidity (%), and daily mean temperature (°C). Grey bars show SD (except for total cover, which is a sum, not a mean). Significant relationships (linear model) shown with a line."}
#' Loop through all combination of dep and indep variables and their models,
#' and make plots for each.
var_grid <- make_var_grid (c("rh_min", "par_total", "temp_mean"))

@@

```{r fig-7, fig.width=7, fig.height=4, fig.cap = "Fig. 7. Sizes of gametophytes with antheridia (male), archegonia (female), or no reproductive structures (asexual). a. Length (mm). b. Width (mm). Boxplots show medians and extend to the first and third quartile; whiskers extend to 1.5× the interquartile range. A single archegoniate (female) outlier with length of 46.6 mm was removed prior to plotting. Points randomly scattered within each x-axis category to reduce overlap."}
#' Note there is one outlier with length 46 mm, 
#' then the rest all less than 19 mm. We will remove this 
#' before plotting and make a note in the figure caption.
gameto_size_no_outlier <-
  gameto_size %>%
  # Change sex to uppercase for nicer plotting
  mutate(sex = stringr::str_to_title(sex)) %>%
  # Remove one outlier
  filter(length < 40)

@@

```{r fig-8, fig.width=7, fig.height=8, fig.cap = "Fig. 8. A comparison of vegetation of the independent gametophyte site with that of a sporophyte site of Pleurosoriopsis. a--b. Typical vegetation of sporophyte habitat covered with summer-green broad-leaved trees (Chichibu-shi, Saitama Pref., Japan). c--d. Vegetation of the independent gametophyte site in Unazawa Valley covered with evergreen cypress trees. a, c. in August, b, d. in December.  "}

magick::image_read_pdf(here("ms/Fig8_vegetation.pdf"))

@@

```{r table-1}
#' Format Table 1 and output to RTF using `gt` package.
t_test_results %>%
  arrange(factor(var, levels = order_vars)) %>%
  mutate(season = stringr::str_to_title(season)) %>%
  mutate(var = case_when(
    var == "rh_min" ~    "Min. Rel. Humidity",
    var == "par_total" ~ "DLI",
    var == "temp_mean" ~ "Mean Temp."
  )) %>%
  group_by(season) %>%
  gt() %>%
  fmt_number(
    columns = vars(estimate, statistic, conf.low, conf.high, 
                   okutama_mean, okutama_sd, 
                   uratakao_mean, uratakao_sd)
  ) %>%
  cols_move(
    columns = vars(okutama_mean, okutama_sd),
    after = vars(conf.high)
  ) %>%
  tab_spanner(
    label = "Okutama",
    columns = vars(okutama_mean, okutama_sd)
  ) %>%
  tab_spanner(
    label = "Uratakao",
    columns = vars(uratakao_mean, uratakao_sd)
  ) %>%
  cols_label(
    var = "",
    statistic = "t",
    pval = "p",
    parameter = "df",
    conf.low = "Lower",
    conf.high = "Upper",
    okutama_mean = "Mean",
    okutama_sd = "SD",
    uratakao_mean = "Mean",
    uratakao_sd = "SD"
  ) %>%
  tab_style(
    style = cells_styles(
      bkgd_color = "white"),
    locations = cells_data(
      columns = everything()
    )
  ) %>%
  tab_header(
    title = "Table 1"
  )

@@

linear_model_results %>%
  select(-fits) %>%
  arrange(factor(x_var, levels = order_vars)) %>%
  mutate(
    x_var = case_when(
      x_var == "rh_min" ~    "Min. Rel. Humidity",
      x_var == "par_total" ~ "DLI",
      x_var == "temp_mean" ~ "Mean Temp."
    ),
    y_var = case_when(
      y_var == "total_cover" ~ "Total cover",
      y_var == "count_mean" ~ "Gemmae count",
      y_var == "length_mean" ~ "Gemmae length"
    )
  ) %>%
  group_by(x_var) %>%
  gt() %>%
  cols_move_to_start(vars(y_var)) %>%
  cols_move(
    columns = vars(df.residual),
    after = vars(df)
  ) %>% 
  fmt_number(
    columns = colnames(.) %in% cols_numeric
  ) %>%
  fmt_number(
    columns = vars(p.value),
    decimals = 3
  ) %>%
  fmt_number(
    columns = vars(deviance),
    decimals = 0
  ) %>%
  cols_label(
    y_var = "",
    r.squared = "r-squared",
    adj.r.squared = "adj. r-squared",
    statistic = "t",
    p.value = "p",
    df.residual = "df residual"
  ) %>%
  tab_style(
    style = cells_styles(
      bkgd_color = "white"),
    locations = cells_data(
      columns = everything()
    )
  ) %>%
  tab_header(
    title = "Table 2"
  )

>>>>

---
title: Growth Dynamics of the Independent Gametophytes of *Pleurosoriopsis makinoi*
  (Polypodiaceae)
output:
  pdf_document:
    latex_engine: xelatex
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)

@@

```{r, results='asis', echo = FALSE, eval=(knitr::opts_knit$get('rmarkdown.pandoc.to') == 'latex')}
# Add pagebreak if rendering to pdf
cat('\\pagebreak')
```

@@

```{r fig-1, fig.width=7, fig.height=7, fig.cap = "The quadrats for measurement of gametophyte cover of *Pleurosoriopsis makinoi* in Unazawa Valley, Okutama-machi, Tokyo, Japan. a. A full view of the site with four quadrat positions (q1, q2, q3 and q4). b. A 10 cm × 10 cm quadrat frame."}
knitr::include_graphics(here("images/Fig1_quadrats.pdf"))

@@

```{r fig-2, fig.width=7, fig.height=7, fig.cap = "Gametophytes of *Pleurosoriopsis makinoi* collected in Unazawa Valley, Okutama-machi, Tokyo, Japan. a. A typical ribbon-like gametophyte with gemmae on the apical part (arrowhead), collected in August. b. Gemmae (arrowheads) on a gametophyte collected in November. c. A gametophyte with archegonia (arrowheads), collected in October. d. Archegonia. e. A gametophyte with antheridia (arrowhead) collected in December. f. An antheridium. Scale bars: 2 mm for a, 200 μm for b and d, 1 mm for c and e, 100 μm for f."}
knitr::include_graphics(here("images/Fig2_gemma_sexualorgans.pdf"))
```

```{r fig-3, fig.width=8, fig.height=7, fig.cap="Change in growth of *Pleurosoriopsis makinoi* gametophytes observed from May 2009 to May 2014. a. Cover (cm^2^) of gametophytes in four 10 × 10 cm quadrats. b. Mean number of gemmae per individual. c. Mean gemma length (μm). For b and c, *n* = 10 individuals per observation; grey bars show SD. Shaded bars show alternating years."}
#' Set x-limits manually since the automatically set limits will vary
#' when we subset the data.
start_date <- combined_morph %>%
  pull(date) %>%
  min

@@

```{r fig-4, fig.width=8, fig.height=7, fig.cap = "Change in microclimate at site with *Pleurosoriopsis makinoi* independent gametophytes (Okutama) and sporophytes (Uratakao) from 2013 to 2014. a--b. Daily light integral (DLI) (mol m^-2 day^-1). c--d. Daily minimum relative humidity (%). e--f. Daily mean temperature (°C). Shaded bars show alternating years. Data collected at Okutama from September 22, 2012 to July 22, 2014 and at Uratakao from August 20, 2013 to November 22, 2014. Some data missing due to equipment malfunction (see Results)."}

@@

```{r fig-5, evaluate = FALSE, fig.width=7, fig.height=8, fig.cap="Density plots showing difference between microclimate at *Pleurosoriopsis makinoi* independent gametophyte site (Okutama) and sporophyte site (Uratakao). Significant differences (paired t-test) shown by asterisks: \\*\\*\\*, *p* < 0.001; \\*\\*, *p* < 0.01; \\*, *p* < 0.05.  Y-axis is relative probability density; curves scaled so that total probability of each scales to 1. Vertical lines shown medians."}
#' Extract p-values for plotting, and put into data frame
#' with xvals specifying season and yvals specifying height
#' to plot asterisks.
p_vals <-
  t_test_results %>%
  mutate(pval = case_when(
    pval < 0.0001 ~ "***",
    pval < 0.001 ~ "**",
    pval < 0.05 ~ "*",
    TRUE ~ ""
  )) %>%
  select(season, var, pval) %>%
  spread(var, pval) %>%
  rename(
    par_total_pval = par_total,
    rh_min_pval = rh_min,
    temp_mean_pval = temp_mean
  )

@@

```{r fig-6, fig.width=7, fig.height=8, fig.cap="Relationships between growth of *Pleurosoriopsis makinoi* gametophytes (total cover, mean gemma number, and mean gemma length) and microclimate at the Okutama site. Microclimatic variables are monthly means of daily light integral (DLI) (mol m^-2^ day^-1^), daily minimum relative humidity (%), and daily mean temperature (°C). Grey bars show SD (except for total cover, which is a sum, not a mean). Significant relationships (linear model) shown with a line."}
#' Loop through all combination of dep and indep variables and their models,
#' and make plots for each.
var_grid <- make_var_grid (c("rh_min", "par_total", "temp_mean"))

@@

```{r fig-7, fig.width=7, fig.height=4, fig.cap = "Sizes of gametophytes with antheridia (male), archegonia (female), or no reproductive structures (asexual). a. Length (mm). b. Width (mm). Boxplots show medians and extend to the first and third quartile; whiskers extend to 1.5× the interquartile range. A single archegoniate (female) outlier with length of 46.6 mm was removed prior to plotting. Points randomly scattered within each x-axis category to reduce overlap."}
#' Note there is one outlier with length 46 mm,
#' then the rest all less than 19 mm. We will remove this
#' before plotting and make a note in the figure caption.
gameto_size_no_outlier <-
  gameto_size %>%
  # Change sex to uppercase for nicer plotting
  mutate(sex = stringr::str_to_title(sex)) %>%
  # Remove one outlier
  filter(length < 40)

@@

```{r fig-8, fig.width=7, fig.height=7, fig.cap = "A comparison of vegetation of the independent gametophyte site with that of a sporophyte site of Pleurosoriopsis. a--b. Typical vegetation of sporophyte habitat covered with summer-green broad-leaved trees (Chichibu-shi, Saitama Pref., Japan). c--d. Vegetation of the independent gametophyte site in Unazawa Valley covered with evergreen cypress trees. a, c. in August, b, d. in December.  "}
knitr::include_graphics(here("images/Fig8_vegetation.pdf"))
```

```{r, results='asis', echo = FALSE, eval=(knitr::opts_knit$get('rmarkdown.pandoc.to') == 'latex')}
# Add pagebreak if rendering to pdf
cat('\\pagebreak')
```

@@

```{r table-1}
#' Format Table 1 and output to RTF using `gt` package.
t_test_results %>%
  arrange(factor(var, levels = order_vars)) %>%
  mutate(season = stringr::str_to_title(season)) %>%
  mutate(var = case_when(
    var == "rh_min" ~    "Min. Rel. Humidity",
    var == "par_total" ~ "DLI",
    var == "temp_mean" ~ "Mean Temp."
  )) %>%
  group_by(season) %>%
  gt() %>%
  fmt_number(
    columns = vars(estimate, statistic, conf.low, conf.high,
                   okutama_mean, okutama_sd,
                   uratakao_mean, uratakao_sd)
  ) %>%
  cols_move(
    columns = vars(okutama_mean, okutama_sd),
    after = vars(conf.high)
  ) %>%
  tab_spanner(
    label = "Okutama",
    columns = vars(okutama_mean, okutama_sd)
  ) %>%
  tab_spanner(
    label = "Uratakao",
    columns = vars(uratakao_mean, uratakao_sd)
  ) %>%
  cols_label(
    var = "",
    statistic = md("*t*"),
    pval = md("*p*"),
    parameter = "df",
    conf.low = "Lower",
    conf.high = "Upper",
    okutama_mean = "Mean",
    okutama_sd = "SD",
    uratakao_mean = "Mean",
    uratakao_sd = "SD"
  ) %>%
  tab_style(
    style = cells_styles(
      bkgd_color = "white"),
    locations = cells_data(
      columns = everything()
    )
  ) %>%
  tab_header(
    title = "Table 1",
    subtitle = md("&nbsp;") # work-around for https://github.com/rstudio/gt/issues/197
  )

@@

linear_model_results %>%
  select(-fits) %>%
  arrange(factor(x_var, levels = order_vars)) %>%
  mutate(
    x_var = case_when(
      x_var == "rh_min" ~    "Min. Rel. Humidity",
      x_var == "par_total" ~ "DLI",
      x_var == "temp_mean" ~ "Mean Temp."
    ),
    y_var = case_when(
      y_var == "total_cover" ~ "Total cover",
      y_var == "count_mean" ~ "Gemmae count",
      y_var == "length_mean" ~ "Gemmae length"
    )
  ) %>%
  group_by(x_var) %>%
  gt() %>%
  cols_move_to_start(vars(y_var)) %>%
  cols_move(
    columns = vars(df.residual),
    after = vars(df)
  ) %>%
  fmt_number(
    columns = colnames(.) %in% cols_numeric
  ) %>%
  fmt_number(
    columns = vars(p.value),
    decimals = 3
  ) %>%
  fmt_number(
    columns = vars(deviance),
    decimals = 0
  ) %>%
  cols_label(
    y_var = "",
    r.squared = md("r2"),
    adj.r.squared = md("adj. r2"),
    statistic = md("*t*"),
    p.value = md("*p*"),
    df.residual = "df resid."
  ) %>%
  tab_style(
    style = cells_styles(
      bkgd_color = "white"),
    locations = cells_data(
      columns = everything()
    )
  ) %>%
  tab_header(
    title = "Table 2",
    subtitle = md("&nbsp;") # work-around for https://github.com/rstudio/gt/issues/197
  )
